---
title: "Sumary of Health Impact - Aircraft UFP"
author: "Magali Blanco"
date: "2024-05-23"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Notes

Health Impact Assumptions
* assuming risk (AF) is the same across populations within an area (i.e., does not differentiate between higher risk groups [e.g., children, elderly, pregnant])


# Setup

```{r setup, include=FALSE}
# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
           detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

pacman::p_load(tidyverse, 
               ggpubr, #ggarrange()
               sf, knitr, kableExtra)

out_path <- file.path("data", "output")

set.seed(1)
 
# ggplot settings
theme_set(theme_bw())
theme_update(legend.position = "bottom")

# run a subset of scenarios
test_run <- FALSE 

##################################################################################################
# DATA
##################################################################################################
load(file.path("data", "modified", "analysis_files.rda"))
load(file.path("data", "output", "health_impact_files.rda"))

health_impact <- health_impact %>% 
  # remove "percent_" from rollback scenario label
  mutate(control_scenario = gsub(".*_", "", control_scenario) %>% as.numeric() %>% as.factor())  
 
if(test_run == TRUE) {
  keep_geoids <- sample(x=unique(pop$GEOID), size = 200, replace = F)
  
  health_impact <- health_impact %>%
    filter(#control_scenario %in% c(25, 50#, 75
                          #         ),
           outcome == first(outcome),
           #grepl("pop_total|male|female|non_hispanic_white|people_of_color", variable),
            !grepl("partner", variable)
           #GEOID %in% keep_geoids
           )
  
}

#unique(health_impact$control_scenario)

##################################################################################################
# FUNCTIONS
##################################################################################################
# function creates pop variable groups
add_pop_variable_group_labels <- function(dt){
  dt %>%
    mutate(
      variable_group = case_when(
        variable == "pop_total" ~ "total",
        grepl("age", variable) ~"age",
        variable %in% c("male", "female") ~"sex",
        grepl("race|hispanic|people_of_color", variable) ~"race_ethnicity",
        grepl("partner|married|spouse", variable) ~"partner",
        grepl("house", variable) ~"housing",
        grepl("poverty|income", variable) ~"poverty",
        
        # --> add others here as needed 
        ))
}

##################################################################################################
# --> function to label 'variable'




##################################################################################################
# function returns distribution
summary_table <- function(dt, variable_name){
  dt %>%
    rename(var = all_of(variable_name)) %>%
    summarize(n=n(),
            Min = min(var),
            Q25 = quantile(var, 0.25),
            Median = median(var),
            Mean = mean(var),
            SD = sd(var),
            Q75 = quantile(var, 0.75),
            Max = max(var)
    )
  }

#################################################################################################
# --> MOVE TO 1_prep....R? DATA SETUP
##################################################################################################
health_impact <- add_pop_variable_group_labels(health_impact) %>%
  # assuming risk is evenly distributed and only based on population counts...
  # --> verify that this methodology is correct 
  mutate(outcome_baseline = pop*incidence)
  
########
control_scenario <- health_impact %>%
  st_drop_geometry() %>%
  distinct(GEOID, #baseline_value = Values, 
           value=Values) %>%
  mutate(scenario="baseline")
  
  
compare_ap <- health_impact %>%
  st_drop_geometry() %>%
  distinct(GEOID, #baseline_value = Values, 
           scenario=control_scenario, value=control_value) %>%
  mutate(scenario = paste0(scenario, "_pct_reduction")) %>%
  bind_rows(control_scenario) %>%
  mutate(scenario = relevel(factor(scenario), ref="baseline"))
  
########
pop <- add_pop_variable_group_labels(pop)  

########
svi <- svi %>%
  mutate(index_variable = case_when(
    index=="svi" ~ gsub("RPL_THEME._", "", index_variable)))

##################################################################################################
# COMMON VARIABLES
##################################################################################################

# --> create groups (age, partner, race-eth...), define levels for e.g., 'variable'


# --> update
map_variables <- c("pop_total", "female", "people_of_color", "age_00_04", "age_65_plus", "partner_opposite_sex", "poverty_below")


variable_groups <- unique(health_impact$variable_group)
outcomes <- unique(health_impact$outcome)

race_levels <- c("non_hispanic_white", "asian", "hispanic", "multiple", "black_aa", "hawaiian_islander", "native", "other")

```

# --> include all ages? doesn't make sense for children? is PTB incidnece for total pop (not just women of childbearing age)?

# --> START HERE: calculate proportion of reducions (subgroup total) for eac subgroup? add 2nd x-axis? 

# Population demographics

* using Census 2020 info

## table counts 

# --> repeat tables for proportions 

```{r}
# table of counts
## demographics
health_impact %>%
  st_drop_geometry() %>%
  distinct(GEOID, variable_group, variable, pop) %>%
  group_by(variable_group, variable) %>%
  summary_table(variable_name = "pop") %>%
  kable(caption = "distribution of total population counts by census tract") %>%
  kable_styling()


# --> vulnerability indeces
svi %>%
  group_by(index, index_variable) %>%
  summary_table(variable_name = "index_value") %>%
  kable(caption = "distribution of vulnerability index value at census tracts") %>%
  kable_styling()


```

## maps

demographics 

# --> repeat for proportions 

```{r}
# this_group = map_variables[1]
lapply(map_variables, function(this_group) {
  tracts %>%
    left_join(filter(pop, variable == this_group), by="GEOID") %>%  
    
    ggplot(aes(fill=pop)) + 
    facet_wrap(~variable) +
    geom_sf() +
    scale_fill_viridis_b(#option="magma"
                          #option="heat"
                         ) +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()
          ) +
    labs(fill="Count")
  
}) %>%
  ggarrange(plotlist = ., nrow = 1)
  
 
```

vulnerability

```{r}
tracts %>%
  left_join(filter(pop, variable == "pop_total"), by="GEOID") %>%
  left_join(svi, by="GEOID") %>%
    
    ggplot(aes(fill=index_value)) + 
    facet_wrap(~index+index_variable) +
    geom_sf() +
    scale_fill_viridis_b() +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()
          ) +
  labs(fill="Count")   

```


# UFP Exposures

## maps 

exposure maps before and after

```{r}
 ##############
left_join(tracts, compare_ap, by="GEOID") %>%
  ggplot(aes(fill=value)) + 
    facet_wrap(~scenario) +
    geom_sf() +
    scale_fill_viridis_c() +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()
          ) +
  labs(fill="UFP (pt/cm3)\nExposure Scenarios")   
  
```

# Incidence at baseline

# --> drop brain tumors (rare outcome)? baseline counts are very low per census tract

overall incidence 

```{r}
# baseline county-level incidence rates
incidence %>%
  kable(caption = "incidence rates applied") %>%
  kable_styling()

# baseline incidence by census tract based on total population
health_impact %>%
  st_drop_geometry() %>%
  filter(variable == "pop_total",
         # only look at mean change in incidence for now
         af_name=="af_mean"
         ) %>%  
  group_by(#af_name, 
           variable,
           outcome, control_scenario) %>%
  summary_table(variable_name="outcome_control") %>%
  kable(caption = "Distribution of the number of cases reduced under control scenarios", digits = 3) %>%
  kable_styling()
  
  

##########################
# map of baseline incidence counts 

print("assuming county-level incidence rates are the same across the population")
print("note, the patterns here are the same across outcomes & the same as the total population map")

lapply(outcomes, function(this_outcome) {
  health_impact %>%
  filter(outcome==this_outcome,
         variable == "pop_total") %>%
  distinct(GEOID, outcome, outcome_baseline, geometry) %>%

  #mutate(outcome_baseline = pop*incidence) %>%
  
  ggplot(aes(fill=outcome_baseline)) + 
  facet_wrap(~outcome) +
    geom_sf() +
    scale_fill_viridis_c() +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()
          ) +
  labs(fill="Baseline Incidence\nSolely Based on\nPopulation Counts")
  }) %>%
  ggarrange(plotlist = .)

```



--> Y: exposure vs X: groups (show exposure disparities & po)

## by groups 

# --> need to do population-weighted exposure estimates for subgroups (e.g., white vs POC)...need the "total"

```{r}
# exposures by group
# tracts %>%
#   left_join(filter(pop, variable %in% map_variables), by="GEOID", relationship="one-to-many") %>%
filter(pop, variable %in% map_variables) %>%
  left_join(compare_ap, by="GEOID", relationship="many-to-many") %>%
  
  
  ggplot(aes(x=variable, y=value, fill=scenario)) +
  facet_wrap(~variable_group, scales="free") + 
  geom_boxplot() + 
  labs(y="")







```



# Health Impact from Air Pollution Reductions

### --> AND/OR look at % reduction from baseline/standardized to pop size (some groups have very small Ns)

## categorical variables

```{r}
# plot1 function


# cases reduced by group
health_impact %>%
  filter(af_name == "af_mean") %>%
  
  ggplot(aes(x=variable, y=outcome_control, fill=control_scenario)) + 
  facet_wrap(~outcome+variable_group, scales="free") + 
  geom_boxplot() +
  labs(y = "No. Cases Reduced")
  
   




 

```


## disadvantage indices

```{r}

# dt_to_merge = svi %>%
#   filter(grepl(#"RPL_THEMES_OVERALL", 
#                "RPL_THEME1_SES",
#                index_variable))

cases_by_continuous_var_plot <- function(dt_to_merge) {
  # --> error: mutliple matches??
health_impact %>%
    # point health estimates
    filter(variable == "pop_total",
           af_name == "af_mean") %>% #filter(GEOID == first(GEOID)) %>% View()
    left_join(dt_to_merge, relationship = "many-to-many", by="GEOID") %>%
    
    ggplot(aes(y=index_value, x=outcome_control, col=index_value, shape=control_scenario, linetype=control_scenario)) +
    facet_wrap(~outcome+index_variable, scales="free_x") +
    geom_point() + 
    geom_smooth() +
    labs(#y= group_name,
       x="Number of Cases Reduced",
       shape = "UFP Reduction (%)",
       linetype = "UFP Reduction (%)",
       )
}

##################################################################################################
svi %>%
  filter(grepl(#"RPL_THEMES_OVERALL", 
               "RPL_THEME1_SES",
               index_variable)) %>%
  cases_by_continuous_var_plot()
  
  


```


## maps

--> map total cases reduced OR proportion of pop w/ case reductions 

```{r}


```



